---
title: "Xh-pi interactions"
author: "Emma Rand"
date: "22/08/2019"
output:
  html_document: bookdown::html_document2
  pdf_document: bookdown::pdf_document2
  word_document: default
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

# Data used
data files in "../xhpi_data_new/" are in three directories, one for each data-set.

* Neu_43_adv neutron diffraction
* Res_43_adv high resolution
* Temp_43_adv high temperature

Number of structures in each varies

Each protein has 3 files

* 7a3h.pdb
* 7a3h.hpml
* 7a3h.pdb.report
.report files are the plain text files containing information about the XH/pi interactions.

Geometric constraints for this data was d<7 and theta<50, although cut-offs of d<4.3 angstroms and theta<25 deg are often used to define XH/pi interactions,  so many of the interactions reported are not actually XH/pi interactions.

```{r pkgs}
library(tidyverse)
library(data.table)
library(cluster)
library(Rtsne)
```

```{r}
# IMPORT HIGH RES
directory <- "../xhpi_data_new/Res_43_adv/"
reports <- list.files(path = directory, pattern = "*.pdb.report")
number <- length(reports)

highres <- do.call("rbind", 
                   lapply(paste0(directory,"/", reports[1:number]),
                          function(x) read.table(x, 
                                                 header = TRUE,
                                                 stringsAsFactors = FALSE)))

# this results in 4305 interactions



# IMPORT NEUTRON DIFFRACTION
directory <- "../xhpi_data_new/Neu_43_adv/"
reports <- list.files(path = directory, pattern = "*.pdb.report")
number <- length(reports)

neutron <- do.call("rbind", 
                   lapply(paste0(directory,"/", reports[1:number]),
                          function(x) read.table(x, 
                                                 header = TRUE,
                                                 stringsAsFactors = FALSE)))

# this results in 772 interactions


# IMPORT HIGH TEMPERATURE
directory <- "../xhpi_data_new/Temp_43_adv/"
reports <- list.files(path = directory, pattern = "*.pdb.report")
number <- length(reports)

hightemp <- do.call("rbind", 
                   lapply(paste0(directory,"/", reports[1:number]),
                          function(x) read.table(x, 
                                                 header = TRUE,
                                                 stringsAsFactors = FALSE)))

# this results in 1788 interactions
```

## Number of interactions
* neutron diffraction: `r nrow(neutron)`
* high resolution:  `r nrow(highres)`
* high temperature:  `r nrow(hightemp)`

```{r}
glimpse(neutron)
```
```{r}
glimpse(highres)
```

```{r}
glimpse(hightemp)
```

Note, we don't have resolution data for highres or hightemp

## Interaction types

Raw counts, not normalised

```{r}
# crosstabulate donor type and acceptor types
# highres
highressum <- highres %>% 
  group_by(x_atom_id, pi_res_id) %>% 
  count() 

ggplot(highressum, aes(x = x_atom_id, y = pi_res_id, fill = n)) +
  geom_tile() +
  ggtitle("High res") + 
  theme(axis.text.x = element_text(angle = 90))

# neutron
neutronsum <- neutron %>% 
  group_by(x_atom_id, pi_res_id) %>% 
  count()

ggplot(neutronsum, aes(x = x_atom_id, y = pi_res_id, fill = n)) +
  geom_tile() +
  ggtitle("Neutron") + 
  theme(axis.text.x = element_text(angle = 90))

# hightemp
hightempsum <- hightemp %>% 
  group_by(x_atom_id, pi_res_id) %>% 
  count()

ggplot(hightempsum, aes(x = x_atom_id, y = pi_res_id, fill = n)) +
  geom_tile() +
  ggtitle("High temp") + 
  theme(axis.text.x = element_text(angle = 90))
```

# Clustering mixed data types

Used R [@R-Core] with packages tidyverse [@tidyverse] and data.table [@data.table] to perform partitioning around medoids (PAM) [@kaufmanl1987clustering, kaufmanl1990clustering] to cluster based on Gower distance [@gower1971] both implemented in the cluster package [@cluster]  


* distance calculation - gower distance
* clustering algorithm - partitioning aroung medoids
* selecting the number of clusters - silhouette width

```{r}

```

## neutron

Some variables will not used in clustering: 
* x_res_num
* x_atom_num
* pdb
* pi_res_num
* resolution

```{r}
# remove variables not used in clustering
neutron2 <- neutron %>%
  select(-x_res_num,
         -x_atom_num,
         -pdb,
         -pi_res_num,
         -resolution)

# char need to be converted to factors
neutron2 <- neutron2 %>% mutate_if(is.character,as.factor)
```

### Distance calculation, no variable transformations applied

```{r}
# calc distances without any transformations in first instance
gower_dist <- daisy(neutron2,
                    metric = "gower")
summary(gower_dist)
```
Which observations are least and most similar?

**most similar i.e., min disimilarity**
```{r}
gower_mat <- as.matrix(gower_dist)
neutron2[which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]),
               arr.ind = TRUE)[1, ], ]
```

**least similar i.e., max disimilarity**
```{r}
neutron2[which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]),
               arr.ind = TRUE)[1, ], ]
```




### Cluster with PAM partitioning round medoids


```{r}
sil_width <- c(NA)
cnum <- 25

for(i in 2:cnum){

  pam_fit <- pam(gower_dist,
                 diss = TRUE,
                 k = i)

  sil_width[i] <- pam_fit$silinfo$avg.width

}
```

# Plot sihouette width (higher is better)
The silhouette width, measure for selecting the number of clusters, was calculated for clusters ranging from 2 to `r cnum`
A higher silhouette width is better. Ideally, one sees it peak and fall again 


```{r}
plot(1:cnum, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:cnum, sil_width)
```


```{r}
sil_width <- c(NA)
cnum <- 80

for(i in 2:cnum){

  pam_fit <- pam(gower_dist,
                 diss = TRUE,
                 k = i)

  sil_width[i] <- pam_fit$silinfo$avg.width

}
```
Since the silhouette width was still increasing, I tried raising to 80.
```{r}
plot(1:cnum, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:cnum, sil_width)
```

<!--  vis -->
<!-- tsne_obj <- Rtsne(gower_dist, is_distance = TRUE) -->

<!-- tsne_data <- tsne_obj$Y %>% -->
<!--   data.frame() %>% -->
<!--   setNames(c("X", "Y")) %>% -->
<!--   mutate(cluster = factor(pam_fit$clustering)) -->

<!-- ggplot(aes(x = X, y = Y), data = tsne_data) + -->
<!--   geom_point(aes(color = cluster)) -->

<!--  consider nature of the variables before clustering -->

<!-- glimpse(neutron2) -->

<!-- hist(neutron2$x_width) -->
<!--  log the bfactors -->
<!-- gower_dist <- daisy(neutron2, -->
<!--                     metric = "gower", -->
<!--                     type = list(logratio = 4, -->
<!--                                 logratio = 8)) -->

<!--  which observations are least and most similar -->
<!--  most similar i.e., min disimilarity -->
<!-- gower_mat <- as.matrix(gower_dist) -->
<!-- neutron2[which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]), -->
<!--                arr.ind = TRUE)[1, ], ] -->

<!--  least similar i.e., max disimilarity -->
<!-- neutron2[which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]), -->
<!--                arr.ind = TRUE)[1, ], ] -->

<!--  cluster with PAM partitioning round medoids -->

<!--  selecting the number of clusters with silhouette width -->
<!--  calculating silhouette width for clusters ranging from 2 to cnum -->
<!--  for the PAM algorithm, we see that 3 clusters yields the highest value. -->
<!-- sil_width <- c(NA) -->
<!-- cnum <- 40 -->

<!-- for(i in 2:cnum){ -->

<!--   pam_fit <- pam(gower_dist, -->
<!--                  diss = TRUE, -->
<!--                  k = i) -->

<!--   sil_width[i] <- pam_fit$silinfo$avg.width -->

<!-- } -->

<!--  Plot sihouette width (higher is better) -->

<!-- plot(1:cnum, sil_width, -->
<!--      xlab = "Number of clusters", -->
<!--      ylab = "Silhouette Width") -->
<!-- lines(1:cnum, sil_width) -->

# references